---
title: "WUP_04"
author: "Emann Malik"
date: "February 27, 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
scores <- read.csv(file = "GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
  
A variable analysis will be conducted for the data set "scores" to determine if there is any difference in Wordsum test scores (WORDSUM) - a numerical variable - among those with varying levels of educational attainment (DEGREE) - a factor variable. We will be using a box plot to view a broad picture of our data and spot any patterns, summary info (mean, standard deviation, etc.) to compare the numerical data of the test scores of each degree level, and a one-way ANOVA test and TukeyHSD plot to determine how many degree levels, if any, are different from each other and thus whether the null or the alternate hypothesis is true.  
  
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
  
```{r}
basicCN <- ggplot(scores,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot(notch=FALSE,outlier.color = "red") + stat_boxplot(geom ='errorbar') + labs(title="Degree vs wordsum score")
```
  
The top of the box in the box plot that corresponds to the LT HIGH SCHOOL Wordsum scores does not touch the bottom of the box of the JUNIOR COLLEGE or the GRADUATE scores. It does touch the bottom of the box for BACHELOR, but just barely. These non-contacts mean that there is some sort of difference between the light high school scores and the junior college, graduate, and bachelor scores. The HIGH SCHOOL box plot's box barely touches the bottom of the JUNIOR COLLEGE box and does not even reach the bottom of the GRADUATE box. It is also relatively far from the median score of the BACHELOR box plot. BACHELOR, GRADUATE, and JUNIOR COLLEGE scores don't seem to differ much from each other since their boxes all overlap, and neither do HIGH SCHOOL and LT HIGH SCHOOL's since their boxes also overlap. The numerical and inferential results will show us these differences and similarities. 
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
  
```{r}
favstats(WORDSUM~DEGREE,data=scores)
```
  
The average junior college score, about 9.92, is very close to the average graduate score, about 10.10. The average bachelor score, about 9.20, is lower than the average junior college score, about 9.92. The average high school (about 8.28) and light high school score (7.75) are the lowest average scores out of the five levels of DEGREE. They are the furthest from the highest average scores out of the five - graduate and junior college. The graduate level had the highest average and median score (10.10 and 11), while the light high school level had the lowest average and median score (7.75 and 8). 
  
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
